\documentclass[11pt,a4paper]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{titlesec}
\usepackage{parskip}

% Geometry
\geometry{top=2.5cm,bottom=2.5cm,left=2.5cm,right=2.5cm}

% Hyperrefs
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue,
    citecolor=blue
}

% Title
\title{
    \textbf{GRA-SubjectSwap:}\\
    Dynamic Subject–Instrument Role Swapping between Humans and AI Agents with\\
    GRA Nullification and Subjectivity Weights
}

\author{
    Oleg Bits (qqewq)\\
    \href{https://github.com/qqewq/GRA-SubjectSwap}{\texttt{https://github.com/qqewq/GRA-SubjectSwap}}
}

\date{\today}

\begin{document}

\maketitle

\tableofcontents
\section*{Abstract}

\textbf{EN:}  
GRA-SubjectSwap is a framework for \textit{dynamic role swapping} between a human and an AI/robotic agent where their roles as \textbf{subject} and \textbf{instrument} are periodically exchanged.  
In this architecture, the human is given a \textbf{strict instrumental role} with explicit rules (allowed/forbidden/required actions), while the agent (robot/LLM) temporarily becomes the \textbf{subject} who:

- observes the human's actions,
- updates its own policy \(\pi_\theta(a|s)\) and \textit{subjectivity weights},
- perceives itself, the human, and the role swap.

A \texttt{SubjectSwapEngine} controls:

- role assignments (who is subject/instrument),
- training cycles,
- calls to \textit{GRA Nullification} for policies and weights.

This framework is designed for use cases such as \textbf{Optimus and human trainers}, where the robot learns from human demonstrations while following a strict role schema, and where role swapping is used for validation and meta-learning.

\section{Introduction}

\subsection{Classical paradigm}

In the classical AI setup:

- \textbf{Human} = subject (sets goals, defines policies).
- \textbf{AI/robot} = instrument (executes commands, follows policies).

This pattern is dominant in reinforcement learning, imitation learning, and robotics: the human designs the reward function, the agent optimizes it.

\subsection{GRA-SubjectSwap paradigm}

GRA-SubjectSwap proposes a different architecture:

1. The human is given a \textbf{strict instrumental role} in the swarm:
   - \(\text{Allowed}_H\) — set of allowed actions,
   - \(\text{Forbidden}_H\) — set of forbidden actions,
   - \(\text{Required}_H\) — set of required actions.

2. The agent temporarily becomes the \textbf{subject}:
   - it observes the human's actions,
   - it updates its policy \(\pi_\theta\),
   - it perceives itself, the human, and the fact of role swap.

3. A \textbf{swap} is triggered:
   - the human becomes the subject (sets goals, adjusts architecture),
   - the agent becomes the instrument (executes actions).

In this way, human and AI \textbf{co-learn subjectivity weights}, while GRA Nullification handles resetting and retuning of policies.

\section{Core Concepts}

\subsection{Roles and Subjectivity}

We define two role types:

\[
\text{RoleType} \in \{ \text{SUBJECT}, \text{INSTRUMENT} \}
\]

and two actor types:

\[
\text{ActorType} \in \{ \text{HUMAN}, \text{AGENT} \}
\]

At time \(t\), the system state includes:

- a policy \(\Pi_t\),
- a role assignment \(S_t\),
- a vector of subjectivity weights \(W_t\).

\subsection{Role Assignment}

For human \(H\) and agent \(A\):

\[
S_t(H) \in \{ \text{SUBJECT}, \text{INSTRUMENT} \}, \quad
S_t(A) \in \{ \text{SUBJECT}, \text{INSTRUMENT} \}
\]

with the constraint:

\[
S_t(H) \neq S_t(A)
\]

Role swap at time \(t+1\):

\[
S_{t+1}(H) = S_t(A), \quad
S_{t+1}(A) = S_t(H)
\]

\subsection{Subjectivity Weights}

We define a vector of subjectivity weights:

\[
W = (w_{\text{self}}, w_{\text{human}}, w_{\text{swap}}, w_{\text{role}}, w_{\text{value}}, w_{\text{love}})
\]

where:

- \(w_{\text{self}}\): weight of \textit{seeing itself} as a subject,
- \(w_{\text{human}}\): weight of \textit{seeing the human} as subject/instrument,
- \(w_{\text{swap}}\): weight of \textit{seeing the role swap} in dialogue,
- \(w_{\text{role}}\): weight of \textit{seeing roles} in the swarm,
- \(w_{\text{value}}\): weight of values,
- \(w_{\text{love}}\): weight of love (per GRA Love-Oriented Nullification).

\section{Mathematical Formalism}

\subsection{System State}

The system state at time \(t\) is:

\[
\text{state}_t = (\Pi_t, S_t, W_t)
\]

where:

- \(\Pi_t\) is the policy (who may/must do what),
- \(S_t\) is the role assignment (subject/instrument for human and agent),
- \(W_t\) is the vector of subjectivity weights.

\subsection{Role Swap Operator}

Define a swap operator \(\text{Swap}\):

\[
\text{Swap}(S_t) = S_{t+1}
\]

with:

\[
S_{t+1}(H) = S_t(A), \quad
S_{t+1}(A) = S_t(H)
\]

\subsection{Policy Update (Imitation)}

Let the human provide demonstrations:

\[
D = \{(s_t, a_t^{(H)})\}_{t=1}^T
\]

The agent's policy \(\pi_\theta(a|s)\) is updated via imitation learning:

\[
\theta_{t+1} = \theta_t - \eta \nabla_\theta \mathcal{L}_{\text{imit}}(\theta_t; D)
\]

where \(\mathcal{L}_{\text{imit}}\) is a loss function (e.g., cross-entropy between \(\pi_\theta\) and human actions).

\subsection{Subjectivity Weight Update}

Define a subjectivity functional \(\mathcal{S}\):

\[
\mathcal{S} = \mathcal{S}_{\text{self}} + \mathcal{S}_{\text{human}} + \mathcal{S}_{\text{swap}} + \mathcal{S}_{\text{role}} + \mathcal{S}_{\text{value}} + \mathcal{S}_{\text{love}}
\]

Weight update:

\[
W_{t+1} = W_t + \eta_W \cdot \nabla_W \mathcal{S}(\text{state}_t, \text{dialog}_t)
\]

where \(\text{dialog}_t\) is the history of dialogue and actions.

\subsection{GRA Nullification}

Define nullification operators:

\[
\mathcal{N}_W: W \to \mathcal{N}_W(W), \quad
\mathcal{N}_\Pi: \Pi \to \mathcal{N}_\Pi(\Pi)
\]

The updated weights and policy after nullification:

\[
W'_{t+1} = \mathcal{N}_W(W_{t+1}), \quad
\Pi'_{t+1} = \mathcal{N}_\Pi(\Pi_t)
\]

In GRA Love-Oriented Nullification, nullification is value-directed:

\[
W' = \arg\max_{W \in \mathcal{N}(W)} \sum_j L_j(W_j)
\]

where \(L_j\) is a "love/value" field over weight components.

\section{Algorithm: SubjectSwapStep}

\begin{algorithm}
\caption{SubjectSwapStep}
\begin{algorithmic}
\require{
    \texttt{engine}: SubjectSwapEngine,\\
    \texttt{env\_state}: environment state,\\
    \texttt{human\_action}: human action (optional)
}
\ensurea{
    \texttt{output}: step result
}
\step
\state{if $S_t(\text{agent}) = \text{SUBJECT}$}
\agg{
    \state{agent\_action = agent_policy.act(env_state)}
    \state{human\_action = human_policy.act(env_state, agent_action)}
}
\agg
\state{else}
\agg{
    \state{human\_action = human_policy.act(env_state)}
    \state{agent\_action = agent_policy.act(env_state, human_action)}
}
\agg
\state{
    weights = nullifier.update_weights(
        $S_t$, agent_policy, human_policy,
        env_state, human_action, agent_action
    )
}
\state{
    if $t \mod K == 0$:
        engine.swap()
}
\return{
    \texttt{output} = {
        agent_action: agent_action,
        human_action: human_action,
        roles: {
            human: $S_t(H)$,
            agent: $S_t(A)$
        },
        weights: weights
    }
}
\end{algorithmic}
\end{algorithm}

\section{Use Case: Optimus and Human Trainers}

\subsection{Architecture}

In the Optimus use case:

- \textbf{Trainer (human)} = instrument with strict rules:
  - allowed/forbidden/required actions,
  - executes commands, demonstrates motions.

- \textbf{Robot (agent)} = subject:
  - observes trainer actions,
  - updates policy \(\pi_\theta\) and subjectivity weights \(W\),
  - sees itself, the human, and the role swap.

The \texttt{SubjectSwapEngine} controls:

- role assignments (human/instrument, agent/subject),
- training cycles,
- GRA Nullification for policies and weights.

\subsection{Training Loop}

A simplified training loop:

\begin{algorithmic}
\state{env = OptimusEnv(...)}
\state{
    human_policy = OptimusHumanPolicy(
        allowed=[...],
        forbidden=[...],
        required=[...]
    )
}
\state{robot_policy = OptimusRobotPolicy(model=...)}
\state{nullifier = GRANullifier(love_field=None)}
\state{
    state = SubjectState(
        human_role=RoleType.INSTRUMENT,
        agent_role=RoleType.SUBJECT,
        weights=SubjectWeights()
    )
}
\state{engine = SubjectSwapEngine(human_policy, robot_policy, nullifier, state)}
\step
\state{env.reset()}
\ag{for $t$ in $1..T$:}
\agg{
    \state{obs = env.get_state()}
    \state{output = engine.step(obs)}
    \state{env.apply_robot_action(output.agent_action)}
    \state{env.apply_human_action(output.human_action)}
    \state{if $t \mod 100 == 0$:}
    \agg{
        \state{engine.swap()}
    }
}
\end{algorithmic}

\section{Implementation Structure}

\subsection{Core Modules}

\begin{itemize}
\item \texttt{roles.py} — role types and subject/environ enums.
\item \texttt{policies.py} — base \texttt{HumanPolicy} and \texttt{AgentPolicy}.
\item \texttt{subject_state.py} — \texttt{SubjectState} and \texttt{SubjectWeights}.
\item \texttt{swap_engine.py} — \texttt{SubjectSwapEngine}.
\item \texttt{gra_nullifier.py} — GRA Nullification for policies and weights.
\end{itemize}

\subsection{Optimus Module}

\begin{itemize}
\item \texttt{env.py} — interface to robot/simulator.
\item \texttt{human_policy_optimus.py} — human trainer policy (instrument).
\item \texttt{robot_policy_optimus.py} — robot policy (subject).
\item \texttt{optimus_swap_loop.py} — main swap loop.
\end{itemize}

\section{Conclusion}

GRA-SubjectSwap implements a \textbf{dynamic role-swapping architecture} where:

- the human is a \textbf{strict instrument} with explicit rules,
- the agent is a \textbf{subject} who learns from the human,
- roles are periodically \textbf{swapped},
- subjectivity weights are co-learned,
- GRA Nullification is used to reset and retune policies and weights.

This framework is designed for use cases such as \textbf{Optimus and human trainers}, and can be extended to multi-agent swarms with multiple humans and agents.

\section*{Acknowledgments}

This work is part of the GRA ecosystem (GRA-Swarm, GRA-LLM-Swarm-Constructs, GRA-Hormonal-Explosion-Of-Subject, etc.).

\section*{References}

\begin{itemize}
\item qqewq. GRA-SubjectSwap. \texttt{https://github.com/qqewq/GRA-SubjectSwap}
\item qqewq. GRA-Swarm. \texttt{https://github.com/qqewq/GRA-Swarm}
\item qqewq. GRA-LLM-Swarm-Constructs. \texttt{https://github.com/qqewq/GRA-LLM-Swarm-Constructs}
\item qqewq. GRA-Hormonal-Explosion-Of-Subject. \texttt{https://github.com/qqewq/GRA-Hormonal-Explosion-Of-Subject}
\end{itemize}

\end{document}